---
title: "Prediction the manner in which the exercise was done"
output: 
  html_document: 
    fig_caption: yes
---
Prediction Assignment Writeup, Practical Machine Learning module on Data Science specialization, *coursera.org*

by *Valeriu Ciorba*  

###Summary
The data come from  [http://groupware.les.inf.puc-rio.br/har](http://groupware.les.inf.puc-rio.br/har). The goal of this project is to use some data from accelerometers in order to predict the manner in which the exercise was done. This is the "*classe*" variable in the *training* data set. Was used *caret package* classification methods: *classification tree algorithm* and *random force*. Was also carried out *3-fold validation* using the *trainControl* function. From the results, it appears that the *random forest model* has the best accuracy for *testing* dataset. Random forest model was applyed to the *testing* dataset in order to predict 20 different *test* cases. 

###Preparing
The datasets have been downloaded to local machine, and saved in the working directory. Below codes loaded the the necessary library and datasets.
```{r, results="hide", message=FALSE}
library(lattice); library(ggplot2); library(caret); library(randomForest); library(rpart);
```

```{r, echo = TRUE, cache=TRUE, message=FALSE}
# download training data file from url
download.file("http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", destfile = "pml-training.csv", method="curl")
# download testing data file from url
download.file("http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", destfile = "pml-testing.csv", method="curl")
# read data
training<-read.csv("pml-training.csv",na.strings=c("NA","#DIV/0!"))
testing<-read.csv("pml-testing.csv",na.strings=c("NA","#DIV/0!"))
```

###Data exploring
```{r, results="show", message=FALSE}
dim(training)
table(training$classe)
```
There are 19622 observation/row in traning dataset, including 160 variables/column. The last column is the target variable *classe*. The most abundant class is **A**.  
There are some variables having a lot of missing values, for simplicity, all variables containing NA values was removed. And also, several variables are not direcly related to the target variable classe. Those varialbes (*“x”, “user_name”*, and all the *time related variables*, such as *“raw_timestamp_part_1”* etc) also was removed.
```{r, results="show", message=FALSE, cache=TRUE}
NA_Count = sapply(1:dim(training)[2],function(x)sum(is.na(training[,x])))
NA_list = which(NA_Count>0)
colnames(training[,c(1:7)])
```
```{r, results="show", message=FALSE, cache=TRUE}
training = training[,-NA_list]
training = training[,-c(1:7)]
training$classe = factor(training$classe)
```

###Modeling with Cross Validation
The problem presenting here is a classification problem. Was used *caret package* classification methods: *classification tree algorithm* and *random force*. Was also carried out *3-fold validation* using the *trainControl* function.
```{r, results="show", message=FALSE, cache=TRUE}
set.seed(1234)
cv3 = trainControl(method="cv",number=3,allowParallel=TRUE,verboseIter=TRUE)
modrf = train(classe~., data=training, method="rf",trControl=cv3)
```

```{r, results="show", message=FALSE, cache=TRUE}
modtree = train(classe~.,data=training,method="rpart",trControl=cv3)
```

The performance of these two model on the *training* dataset was:
```{r, results="show", message=FALSE, cache=TRUE}
prf=predict(modrf,training)
ptree=predict(modtree,training)
table(prf,training$classe)
```

```{r, results="show", message=FALSE, cache=TRUE}
table(ptree,training$classe)
```

The performance of these two model on the *testing* dataset:
```{r, results="show", message=FALSE, cache=TRUE}
prf=predict(modrf,testing)
ptree=predict(modtree,testing)
table(prf,ptree)
```
From the results, it appears that the *random forest model* has the best accuracy for *testing* dataset.

### Conclusion
Random forest model was applyed to the *testing* dataset in order to predict 20 different *test* cases. 
```{r, results="show", message=FALSE, cache=TRUE}
answers=predict(modrf,testing)
answers
```
Using recomended function *pml_write_files* to create the files for Prediction Assignment Submission.
```{r, results="show", message=FALSE, cache=TRUE}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
pml_write_files(answers)
```
The predicted classes for the 20 tests are: B A B A A E D B A A B C B A E E A B B B.